---
title: 'Probabilities in R'
author: "Singh Karanbir"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Simulate rolling two dice and adding their values. Perform 10000 simulations and make a bar chart showing how many of each outcome occurred.

```{r}
die1 <- sample(x = 1:6, size = 10000, replace = TRUE)
die2 <- sample(x = 1:6, size = 10000, replace = TRUE)
sumDice <- die1 + die2
head(sumDice)
barplot(table(sumDice))
```

You can buy trick dice, which look (sort of) like normal dice. One die has numbers 5, 5, 5, 5, 5, 5. The other has numbers 2, 2, 2, 6, 6, 6. Simulate rolling the two trick dice and adding their values. Perform 10000 simulations and make a bar chart showing how many of each outcome occurred.
```{r}
die1 <- sample(c(5,5,5,5,5,5), size = 10000, replace = TRUE)
die2 <- sample(c(2,2,2,6,6,6), size = 10000, replace = TRUE)
sumDice <- die1 + die2
head(sumDice)
barplot(table(sumDice))
```

Sicherman dice also look like normal dice, but have unusual numbers. One die has numbers 1, 2, 2, 3, 3, 4. The other has numbers 1, 3, 4, 5, 6, 8. Simulate rolling the two Sicherman dice and adding their values. Perform 10000 simulations and make a bar chart showing how many of each outcome occurred. How does your answer compare to part a?


```{r}
die1 <- sample(c(1,3,4,5,6,8), size = 10000, replace = TRUE)
die2 <- sample(c(1,2,2,3,3,4), size = 10000, replace = TRUE)
sumDice <- die1 + die2
head(sumDice)
barplot(table(sumDice))
```


Two dice are rolled.

What is the probability that the sum of the numbers is exactly 10?

Sample Space is 36 possible outcomes.

Sum == 10 only when -> (5,5),(6,4)(4,6) = 3/36 = 1/12

```{r}


eventA <- replicate(10000, {
  dieRoll <- sample(1:6, 2, replace = TRUE)
  (sum(dieRoll) == 10)
})
probA <- mean(eventA)
probA

```

What is the probability that the sum of the numbers is at least 10?

Sum >= 10 -> (5,5),(6,4),(4,6),(6,6)(5,6)(6,5) = 6/36 = 1/6

```{r}


eventB <- replicate(10000, {
  dieRoll <- sample(1:6, 2, replace = TRUE)
  (sum(dieRoll) >= 10)
})
probB <- mean(eventB)
probB

```

What is the probability that the sum of the numbers is exactly 10, given that it is at least 10?

P(x>=0 given that x == 10) = P(x == 10)/P(x>=0) = (3/36)/(6/36) = 1/2

```{r}


probC <- probA/probB
probC


```


A hat contains six slips of paper with the numbers 1 through 6 written on them. Two slips of paper are drawn from the hat (without replacing), and the sum of the numbers is computed.


What is the probability that the sum of the numbers is exactly 10?

Sample space reduces to 30 since we cannot have (n,n) and n = 6, which out of 36 possible outcomes reduce to 30.

Sum == 10 -> (4,6)(6,4) = 2/30 = 1/15

```{r}
eventD <- replicate(10000, {
  slip <- sample(1:6, 2, replace = FALSE)
  (sum(slip) == 10)
})
probD <- mean(eventD)
probD
```

What is the probability that the sum of the numbers is at least 10?

Sum >= 10 -> (6,4),(4,6),(6,5)(5,6) = 4/30 = 2/15

```{r}
eventE <- replicate(10000, {
  slip <- sample(1:6, 2, replace = FALSE)
  (sum(slip) >= 10)
})
probE <- mean(eventE)
probE

```

What is the probability that the sum of the numbers is exactly 10, given that it is at least 10?

P(x>=0 given that x == 10) = P(x == 10)/P(x>=0) = (1/15)/(2/30) = 1/2

```{r}
probF <- probD/probE
probF
```
